{\bf Demystifying Debugging}

### 2012-11-06 gpi-debug/debugging.tex
add time diagrams and 213 disclaimer

\institute[98-172]{Great Practical Ideas for Computer Scientists}
\date[]{2012, November 1}

@@

\begin{frame}{Asking for Help}
	This DOESN'T mean...
	\begin{itemize}
		\item That TAs can't help you without a detailed story
		\item (That you shouldn't seek help until you've already found your bug?!)
	\end{itemize}
	\pause
	This does mean...
	\begin{itemize}
		\item You need a detailed story to understand your bug.
		\item If the TA doesn't know the story, how can they help?
		\pause
		\begin{itemize}
			\item They know what you should do to make your story better.
		\end{itemize}
	\end{itemize}
\end{frame}

@@

	\linegap
	Here is a bad test case. How can we improve it?

	\linegap
		\texttt{~~~~test\_insert\_sorted()~\{} \\
		\texttt{~~~~~~~~input~=~\hilight{brickred}{[99,~42,~3,~100,~87,~56,~22]};} \\
		\texttt{~~~~~~~~output~=~\hilight{brickred}{[]};} \\
		\texttt{~~~~~~~~\hilight{brown}{foreach}~element~\hilight{brown}{in}~input~\{} \\
		\texttt{~~~~~~~~~~~~insert\_sorted(output,~element);} \\
		\texttt{~~~~~~~~\}} \\
		\texttt{~~~~~~~~\hilight{brown}{return}~is\_sorted(output);} \\
		\texttt{~~~~\}} \\
\end{frame}

@@

	\linegap
	What does this mean 
	% Text: "First I ... then I ... but then it ...!"
\end{frame}

\begin{frame}{Print debugging}
	% Insert some code: the test case with print statements inserted in it
	% Insert a picture: the time-cloud with some printed-out things in it
	Should usually be the first thing you try!
\end{frame}

@@

>>>>

\institute[98-172]{Great Practical Ideas for Computer Scientists}
\date[]{2012, November 8}

@@

\begin{frame}{Aside: Asking for Help}
	This DOESN'T mean...
	\begin{itemize}
		\item That TAs can't help you without a detailed story
		\item (That you shouldn't seek help until you've already found your bug?!)
	\end{itemize}
	\pause
	This does mean...
	\begin{itemize}
		\item You need a detailed story to understand your bug.
		\item If the TA doesn't know the story, how can they help?
		\pause
		\begin{itemize}
			\item They know what you should do to make your story better.
		\end{itemize}
	\end{itemize}
\end{frame}

@@

	\linegap
	Consider this test case:

	\linegap
		\texttt{~~~~input~=~\hilight{brickred}{[99,~42,~3,~100,~87,~56,~22]};} \\
		\texttt{~~~~output~=~\hilight{brickred}{[]};} \\
		\texttt{~~~~\hilight{brown}{foreach}~element~\hilight{brown}{in}~input~\{} \\
		\texttt{~~~~~~~~insert\_sorted(output,~element);} \\
		\texttt{~~~~\}} \\
		\texttt{~~~~\hilight{brown}{return}~is\_sorted(output);~\hilight{darkcyan}{//~returns~false~:(}} \\
	\linegap

	\pause
	\begin{itemize}
		\item Says there is a bug
		\item Doesn't say where
	\end{itemize}
	\pause
	{\bf How can we improve it?}
\end{frame}

@@

	\linegap
	{\em Time-wise}, this means ``It did X, then should have done Y but did Z''
	\begin{itemize}
		\item Are X,Y,Z in enough detail?
		\item Did something important happen in between?
	\end{itemize}
\end{frame}

\begin{frame}{Telling a Story of Time}
	Our test case tells this story:

	\linegap
	\includegraphics[width=\textwidth]{time0.png}
\end{frame}

\begin{frame}{Print Debugging}
	Make your program print its state during the big pink ``???''.
	\begin{itemize}
		\item Reduce how much of the story the ``???'' obscures.
		\begin{itemize}
			\item Find the {\em last} print {\em before} the bug
			\item Find the {\em first} print {\em after} the bug
		\end{itemize}
	\end{itemize}
	\pause

	\linegap
		\texttt{~~~~\hilight{brown}{foreach}~element~\hilight{brown}{in}~input~\{} \\
		\texttt{~~~~~~~~\hilight{olivegreen}{print(output);}} \\
		\texttt{~~~~~~~~insert\_sorted(output,~element);} \\
		\texttt{~~~~\}} \\
\end{frame}

\begin{frame}{Print Debugging}
	New story:

	\linegap
	\includegraphics[width=\textwidth]{time1.png}
\end{frame}

@@

\begin{frame}{Aside: 15-213 ``malloc lab''}
	At the end of 15-213, you will implement \texttt{malloc()}.
	\begin{itemize}
		\item Complex, low-level systems code
		\item Very easy to end up with difficult, subtle bugs
		\begin{itemize}
			\item Your data winds up in the wrong place, gets clobbered by client code
			\item Stale pointers that point to nonsense memory
		\end{itemize}
		\item Your code won't even misbehave immediately, only ``later''.
	\end{itemize}
	\pause
	\linegap

	The assignment will ask you to write a {\bf heap checker}.
	\begin{itemize}
		\item In essence, a very powerful \texttt{assert} for your code.
		\item Debugging and asserts are friends!
	\end{itemize}
\end{frame}

\begin{frame}{Aside: 15-213 ``malloc lab''}
	Many students think: ``I just want to write \texttt{malloc()}. I can do the heap checker later.''
	\pause
	\begin{itemize}
		\item Think of it like a C0 contract that must hold forever.
		\item \textbf{Write the heap checker simultaneously!}
			\pause
			\begin{itemize}
				\item Save time/effort debugging everything else
				\item Your debugging skill will level up
				\item Ben jumps up and down for emphasis
			\end{itemize}
	\end{itemize}
\end{frame}
